---
title: "Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries loading

```{r echo=FALSE}

library(recommenderlab)
library(ggplot2)                       
library(data.table)
library(reshape2)
library(data.table)
library(dplyr)
library(tidyverse)
library(readxl)
library(janitor)
library(mosaic)
library(ggthemes)
library(lubridate)
library(here)
library(skimr)
library(vroom)
library(readr)
library(huxtable)
library(ggpubr)
library(GGally)


```


## Data loading and wrangling

```{r pressure, echo=FALSE}

# Load Data Sets for Income

male_income <- read_excel("Estimated gross income male.xlsx")
female_income <- read_excel("Estimated gross income female.xlsx")

# Due to data_errors, the HDI CSV has to be edited manually before it could be imported.
# In the original file 2 different tables were included and the second one for regions and high/medium/low definition had to be erased
# Furthermore, there were changing delimiters between data points (double quote, double comma,...) and ".." for NA
# Finally, the values for 2017 and 2018 were not adjusted and too high by the factor of 1000

hdi <- vroom("HDI cleaned.csv", skip = 1)

```

```{r}

# Take Initial Look at Data

glimpse(male_income)
glimpse(female_income)
glimpse(hdi)
head(male_income,5)
head(female_income,5)
head(hdi,5)

# clean names for hdi

names(hdi)[names(hdi) == "HDI Rank (2018)"] <- "hdi_rank_2018"

# Set column type right for hdi

hdi <- hdi %>% 
  mutate("1990" = as.numeric("1990"))

# Select only the years used in the income data frame

hdi <- hdi %>% 
  select(Country,
         "1995","2000","2005","2010",
         "2011","2012","2013","2014",
         "2015","2016","2017","2018")

# Pivot longer hdi the table to get years as values in a column
hdi <- hdi %>% 
  pivot_longer(2:13, 
               names_to = "Year", 
               values_to = "hdi")

# Pivot longer both the table to get years as values in a column
male_income <- male_income %>% 
  pivot_longer(3:14, 
               names_to = "Year", 
               values_to = "income_male")

female_income <- female_income %>% 
  pivot_longer(3:14, 
               names_to = "Year", 
               values_to = "income_female")

# inner join male and female tables on country
all_income <- merge(male_income, 
                    female_income, 
                    by = c("Country",
                           "Year","HDI Rank (2018)"))

# Join hdi on income data frame
all_income_hdi <- merge(all_income,
                        hdi,
                        by = c("Year","Country"))


# Look at the first rows
head(all_income_hdi)

# Export as CSV file "Income Data"
write.csv(all_income_hdi, 
          "all_income_hdi.csv")

# Clean names using Janitor package
all_income_hdi <- clean_names(all_income_hdi)

# Create a continent variable
library(countrycode)

continent <- countrycode(sourcevar = all_income_hdi[, "country"],
                            origin = "country.name",
                            destination = "continent")

all_income_hdi <- cbind(all_income_hdi, continent)

# create a gender pay gap (%) variable as percentage of male income

all_income_hdi <- all_income_hdi %>% 
  mutate(perc_gender_pay_gap = ((income_male - income_female)/income_male)*100)
str(all_income_hdi)

#calculate summary statistics and observe missing values if any
skimr::skim(all_income_hdi)

```

## Data cleaning
```{r}

# Check for impossible or unrealistic data in HDI

all_income_hdi %>% 
  filter(hdi > 1)

# None found

# Fix data types and delete empty rows
all_income_hdi <- all_income_hdi %>% 
  filter(!is.na(all_income_hdi)) %>%   #delete empty rows
  mutate(year = as.double(year), # year as a numerical
         continent = factor(continent))

str(all_income_hdi)

#calculate summary statistics and observe missing values if any
skimr::skim(all_income_hdi)


```

### Visualise Extreme Data 

In this part we are visualising our dataset using boxplots, in order to identify outliers and proceed with filtering our extreme values.

```{r}

# Observe extreme values

cleaning_box_1 <- 
  ggplot(all_income_hdi, aes(y=hdi_rank_2018)) +
  geom_boxplot(colour="#00A699")+
  labs(y="HDI Rank")+
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68")) + 
  NULL

cleaning_box_2 <-
  ggplot(all_income_hdi, aes(y=income_male)) +
  geom_boxplot(colour="#00A699")+
  labs(y="Male Income")+
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68")) + 
  NULL

cleaning_box_3 <-
  ggplot(all_income_hdi, aes(y=income_female)) +
  geom_boxplot(colour="#00A699")+
  labs(y="Female Income") +
  theme(plot.title = element_text(face = "bold")) +
  theme(plot.subtitle = element_text(face = "bold", color = "grey35")) +
  theme(plot.caption = element_text(color = "grey68")) + 
  NULL


combined_box_plots <- 
  ggarrange(cleaning_box_1, 
            cleaning_box_2, 
            cleaning_box_3,
            ncol = 2, nrow = 2) +
  theme_bw()+
  theme(panel.border = element_rect(fill=NA, size=1))
  NULL

annotate_figure(combined_box_plots, top = text_grob("Extreme Values by Variable", face = "bold"))

# Explore the extreme values to ensure they make sense
all_income_hdi %>% 
  filter(income_male > 100000 | income_female > 60000) %>% 
  arrange(desc(income_female))

## As expected the middle east mainly accounts for the extreme values in male income, whilst interestingly enough Singapore and Brunei account for the majority of extreme values in female incomes. Nevertheless, following research we can conclude that there are no data entry issues with our dataset

# visualise the dependent variable


```

## Data Exploration
```{r}
# Find the top 10 highest paygap and the lower 10 paygap for 2018
top_10 <- all_income_hdi %>% 
  filter(year == 2018) %>% 
  top_n(10, perc_gender_pay_gap)

low_10 <- all_income_hdi %>% 
  filter(year == 2018) %>% 
  top_n(-10, perc_gender_pay_gap)

# Find average pay gap in 2018 by continent
by_continent <- all_income_hdi %>% 
  filter(year == 2018) %>% 
  group_by(continent) %>%
  summarise(avg_perc_paygap = mean(perc_gender_pay_gap),
            avg_hdi = mean(hdi)) %>% 
  arrange(desc(avg_hdi))

# Find the number of countries with a paygap greater than 50% by continent in 1995 vs 2018
change <- all_income_hdi %>% 
  filter(year %in% c(1995, 2018)) %>% 
  group_by(continent, year) %>% 
  summarise(count = count(perc_gender_pay_gap > 50))
## Things are getting better!!!

```

## Correlation Matrix 

```{r}


#create correlation matrix for variables with data type "double"
correlation_matrix <- all_income_hdi %>% 
  summarise(
         year,
         hdi,
         income_male,
         income_female,
         continent,
         perc_gender_pay_gap)

#have a look on the correlation matrix data
glimpse(correlation_matrix)


#show the correlations
ggpairs(correlation_matrix, columns=1:6,
    upper=list(continuous='cor'),
    lower=list(continuous = 'points'),
    diag=list(continuous='density'),
    axisLabels='show') +
    theme_bw()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
  
```
### Hypothesis Testing 

### Regression Analysis

Our first models are shown below:
```{r}

##makes first 2 models using instructions from Kostis
model1 <- lm(perc_gender_pay_gap ~
               hdi,
             data = all_income_hdi)

model2 <- lm(perc_gender_pay_gap ~
               hdi +
             year,
             data = all_income_hdi)
#model 3 adds the factor variable continent
model3 <- lm(perc_gender_pay_gap ~
               hdi +
             year +
             continent,
             data = all_income_hdi)

huxreg(model1, model2, model3)

str(all_income_hdi)
## still trying to find more variables to utilise to make a more meaningful model 
```

### Visualisation Explainations
1. Global Gender Inequality 
2. Gender Inequality within EU
3. Dimension 1: Health -> Female reproductive health index
4. Dimension 2: Empowerment -> Female empowerment index
5. Dimension 3: Labour -> Female labor market index
